---
title: "GPR68 Deep Mutational Scan replicate corelations"
author: "Matthew Howard, UCSF"
date: "March 11, 2024"
output: html_notebook
---
this notebook contains code used to analyze replicates scores from the GPR68 mutational scan dataset

screening conditions: pH 6.5, pH 5.5, and surface expression

abbreviations and general naming conventions:
decimals are omitted when referring to pH conditions, example:"ph55" for pH 5.5

set working directory, reset to source folder on your system 
```{r}
setwd("/Users/mkh/Library/CloudStorage/Box-Box/WCM Lab/Matt/Lab/Projects/proton_sensors/GPR68_DMS/figures") 
```

source R packages
```{r}
install.packages("tidyverse")
install.packages("gridExtra")
install.packages("ggpubr")
library(tidyverse)
library(gridExtra)
library(ggpubr)
```

pull replicate scores from tsv file
```{r}
replicates<-read_tsv('GPR68_replicate_scores.tsv')
replicates <- na.omit(replicates)
```
plot replicate correlations as scatters for each condition in a grid, r1 vs r2, r1 vs r3, r2 vs r3
```{r}
a <- ggscatter(replicates, x = "ph_5.5_R1", y = "ph_5.5_R2", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-6,6) + ylim(-5,10)
b <- ggscatter(replicates, x = "ph_5.5_R1", y = "ph_5.5_R3", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-6,6) + ylim(-5,10)
c <- ggscatter(replicates, x = "ph_5.5_R2", y = "ph_5.5_R3", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-6,6) + ylim(-5,10)

d <- ggscatter(replicates, x = "ph_6.5_R1", y = "ph_6.5_R2", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-5,10) + ylim(-5,12)
e <- ggscatter(replicates, x = "ph_6.5_R1", y = "ph_6.5_R3", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-5,10) + ylim(-5,12)
f <- ggscatter(replicates, x = "ph_6.5_R2", y = "ph_6.5_R3", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-5,10) + ylim(-5,12)

g <- ggscatter(replicates, x = "surface_R1", y = "surface_R2", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-6,6) + ylim(-6,10)
h <- ggscatter(replicates, x = "surface_R1", y = "surface_R3", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-6,6) + ylim(-6,10)
i <- ggscatter(replicates, x = "surface_R2", y = "surface_R3", 
          cor.coef = TRUE, cor.method = "spearman", size = 0.25, cor.coef.size = 2) +
  theme(text = element_text(size = 6)) +
  xlim(-6,6) + ylim(-6,10)

overall_correlations <- grid.arrange(a,b,c,d,e,f,g,h,i, ncol=3, nrow=3)
ggsave("supplemental/supp_figure16/replicate_correlations.png", overall_correlations, width = 100, height = 100, units = "mm", dpi = 450)

```