---
title: "GPR68 Deep Mutational Scan mean score snapshots"
author: "Matthew Howard, UCSF"
date: "March 11, 2024"
output: html_notebook
---
this notebook was used to analyze mutational tolerance for GPR68 data

abbreviations and general naming conventions:
decimals are omitted when referring to pH conditions, example:"ph55" for pH 5.5

set working directory, reset to source folder on your system 
```{r}
setwd("/Users/mkh/Library/CloudStorage/Box-Box/WCM Lab/Matt/Lab/Projects/proton_sensors/GPR68_DMS/figures") 
```

packages
```{r}
install.packages("ggsci")
install.packages("zoo")
install.packages("tidyverse")
install.packages("patchwork")
library(patchwork)
library(tidyverse)
library(ggsci)
library(zoo)
```

load variant scores file, parse into missense df
```{r}
gpr68_scores<-read.csv('gpr68_scores_processed.csv')

colnames(gpr68_scores) <- c("hgvs", "ph55_SE", "ph55_epsilon", "ph55_score", "ph65_SE", "ph65_epsilon", "ph65_score", "surf_SE", "surf_epsilon", "surf_score",'pos','len','mutation_type','variants','is.wt')

order_missense <- c("A", "G", "M", "V", "L", "I", "T", "S", "C", "Q", "N", "Y", "W", "F", "E", "D", "H", "K", "R", "P")

gpr68_scores_missense <- gpr68_scores %>% filter(variants %in% order_missense)
```

Calculate positional average score (mutational tolerance)
```{r}
gpr68_missense_mut_tolerance <- gpr68_scores_missense %>% 
  group_by(pos) %>% 
  summarise(
    ph55_mean = mean(ph55_score, na.rm=TRUE),
    ph65_mean = mean(ph65_score, na.rm=TRUE), 
    surface_mean = mean(surf_score, na.rm=TRUE),
    ph55_sd = sd(ph55_score, na.rm=TRUE),
    ph65_sd = sd(ph65_score, na.rm=TRUE),
    surf_sd = sd(surf_score, na.rm=TRUE),
    ph55_var = var(ph55_score, na.rm=TRUE),
    ph65_var = var(ph65_score, na.rm=TRUE),
    surf_var = var(surf_score, na.rm=TRUE)
  )  %>%
  mutate(
    ph55_rollmean = rollmean(ph55_mean, 5, na.pad=TRUE),
    ph65_rollmean = rollmean(ph65_mean, 5, na.pad=TRUE),
    surf_rollmean = rollmean(surface_mean, 5, na.pad=TRUE),
    ph55_sd_rollmean = rollmean(ph55_sd, 5, na.pad=TRUE),
    ph65_sd_rollmean = rollmean(ph65_sd, 5, na.pad=TRUE),
    surf_sd_rollmean = rollmean(surf_sd, 5, na.pad=TRUE),
    ph55_var_rollmean = rollmean(ph55_var, 5, na.pad=TRUE),
    ph65_var_rollmean = rollmean(ph65_var, 5, na.pad=TRUE),
    surf_var_rollmean = rollmean(surf_var, 5, na.pad=TRUE)
  )
```

mutational tolerance vs position plots ph 5.5 and ph 6.5
```{r}
# Set the font, size, blah, blah
custom_theme <- theme(
  text = element_text(family = "sans", size = 6, color = "black"),
  panel.border = element_rect(size = 1),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.text = element_text(family = "sans", size = 6, color = "black"),  
  axis.title = element_text(family = "sans", size = 6, color = "black")#,
  #aspect.ratio = 33/100
)

#plot together
ph_means <- ggplot(gpr68_missense_mut_tolerance, aes(x = pos)) + 
  annotate("rect", xmin = 19, xmax = 50, ymin = -Inf, ymax = Inf, alpha = 0.5, fill = "lightgrey") +
  annotate("rect", xmin = 56, xmax = 82, ymin = -Inf, ymax = Inf, alpha = 0.5, fill = "lightgrey") +
  annotate("rect", xmin = 92, xmax = 123, ymin = -Inf, ymax = Inf, alpha = 0.5, fill = "lightgrey") +
  annotate("rect", xmin = 136, xmax = 157, ymin = -Inf, ymax = Inf, alpha = 0.5, fill = "lightgrey") +
  annotate("rect", xmin = 181, xmax = 213, ymin = -Inf, ymax = Inf, alpha = 0.5, fill = "lightgrey") +
  annotate("rect", xmin = 221, xmax = 251, ymin = -Inf, ymax = Inf, alpha = 0.5, fill = "lightgrey") +
  annotate("rect", xmin = 259, xmax = 287, ymin = -Inf, ymax = Inf, alpha = 0.5, fill = "lightgrey") +
  geom_line(aes(y = ph55_rollmean), linetype = "solid", size = 0.75, color = "BLUE") + 
  geom_line(aes(y = ph65_rollmean), linetype = "solid", size = 0.75, color = "RED") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  labs(
    #title = "GPR68 Mutational Tolerance",
    x = "Position (AA)",
    y = "Mean Positional Score"
  ) +
  theme_minimal() +
  theme_bw() +
  scale_x_continuous(
    breaks = seq(0, 365, 50),
    #expand = c(0, 0),
    limits = c(0, 365)
  ) +
  scale_y_continuous(
    #expand = c(0, 0),
    limits = c(-2.0, 1.0)
  ) +
  custom_theme +  # Apply custom font, border size, centered title, and legend
  scale_color_manual(
    values = color_palette,
    labels = c(
      "ph55_rollmean" = "pH 5.5",
      "ph65_rollmean" = "pH 6.5"
    )
  ) +
  guides(color = guide_legend(
    ncol = 2,
    title = "pH Levels",
    label.theme = element_text(size = 6)
  ))

ph_means
```   

make violin plots for the groups of ph55 scores
```{r}
gpr68_scores$group <- "c_Not designated"

gpr68_scores$group <- ifelse(gpr68_scores$pos %in% c(1:18),"b_N-term",
                             ifelse(gpr68_scores$pos %in% c(18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 113, 114, 115, 116, 117, 118, 119, 120, 121, 122, 123, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147, 148, 149, 150, 151, 152, 153, 154, 155, 156, 157, 181, 182, 183, 184, 185, 186, 187, 188, 189, 190, 191, 192, 193, 194, 195, 196, 197, 198, 199, 200, 201, 202, 203, 204, 205, 206, 207, 208, 209, 210, 211, 212, 213, 221, 222, 223, 224, 225, 226, 227, 228, 229, 230, 231, 232, 233, 234, 235, 236, 237, 238, 239, 240, 241, 242, 243, 244, 245, 246, 247, 248, 249, 250, 251, 259, 260, 261, 262, 263, 264, 265, 266, 267, 268, 269, 270, 271, 272, 273, 274, 275, 276, 277, 278, 279, 280, 281, 282, 283, 284, 285, 286, 287),"e_Transmembrane",
                                    ifelse(gpr68_scores$pos %in% c(300:365),"a_C-term",
                                           ifelse(gpr68_scores$pos %in% c(258, 13, 94, 172),"f_Disulfide",
                                                  ifelse(gpr68_scores$pos %in% c(54, 55, 56, 60, 
                                                                     118, 119, 122, 123, 126, 127, 
                                                                     130, 133, 
                                                                     205, 208, 211, 212, 
                                                                     215, 216, 217, 218, 219, 
                                                                     222, 
                                                                     225, 226, 
                                                                     229, 230, 
                                                                     286, 287, 289, 290, 291, 292),"d_G-protein",
                                                                gpr68_scores$group)))))
    
violin_labels <- c("C-term", "N-Term", "Not designated", "G-protein", "Transmembrane", "Disulfide")

ph55_violins <- ggplot(gpr68_scores, aes(x = group, y = ph55_score, color = group)) + 
  geom_hline(yintercept = 0, linetype = "dotted", color = "black") +
  geom_violin(aes(group = reorder(group, ph55_score, mean))) +
  geom_boxplot(width = 0.2, color = "black", outlier.size = 0.5) +
  scale_x_discrete(labels= violin_labels) +
  theme_minimal() +
  labs(x = "",
       y = "pH 5.5 cAMP Signaling Score") +
  theme(legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        text = element_text(size = 6, family = "sans"),
        axis.text = element_text(color = "black")) +
  scale_y_continuous(breaks = c(-2.5, 0, 2.5))

ph55_violins
```

# compile plot 
```{r}
ph55_plots <- ph_means / ph55_violins  + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 6, face = "bold"))

ph55_plots

ggsave("supplemental/mutational_tolerance/ph55_plots.png", ph55_plots, width = 100, height = 75, units = "mm", dpi = 450)
```

