---
title: "GPR68 Deep Mutational Scan Density Plots"
author: "Matthew Howard, UCSF"
date: "March 11, 2024"
output: html_notebook
---
this notebook contains code used to create density plots for GPR68 deep mutational scanning data

screening conditions: pH 6.5, pH 5.5, and surface expression

abbreviations and general naming conventions:
decimals are omitted when referring to pH conditions, example:"ph55" for pH 5.5

set working directory, reset to source folder on your system 
```{r}
setwd("/Users/mkh/Library/CloudStorage/Box-Box/WCM Lab/Matt/Lab/Projects/proton_sensors/GPR68_DMS/figures") 
```

source R packages
```{r}
install.packages("tidyverse")
install.packages("ggsci")
install.packages("patchwork")
library(tidyverse)
library(ggsci)
library(patchwork)
```

pull in DMS data
```{r}
gpr68_scores<-read.csv('gpr68_scores_processed.csv')

colnames(gpr68_scores) <- c("hgvs", "ph55_SE", "ph55_epsilon", "ph55_score", "ph65_SE", "ph65_epsilon", "ph65_score", "surf_SE", "surf_epsilon", "surf_score",'pos','len','mutation_type','variants','is.wt')

# rename cells with insertions and deletions "indels"
gpr68_scores$mutation_type[gpr68_scores$mutation_type == "I"] <- "Indel"
gpr68_scores$mutation_type[gpr68_scores$mutation_type == "D"] <- "Indel"
gpr68_scores$mutation_type[gpr68_scores$mutation_type == "S"] <- "Syn"
gpr68_scores$mutation_type[gpr68_scores$mutation_type == "M"] <- "Mis"
```

ph signaling assay score density figures
```{r}
# define theme elements
my_theme <- theme_minimal() +
    theme(
      legend.position = "bottom",
      text = element_text(size = 10, color = "black", family = "sans"),
      panel.border = element_rect(color = "black", fill=NA, size=1),
      plot.title = element_text(hjust = 0.5),
      panel.grid.major = element_blank(),  
      panel.grid.minor = element_blank(),
      axis.text = element_text(size = 10, color = "black", family = "sans"),
      axis.ticks.x = element_line(size = 1, colour = "black"), 
      axis.ticks.y = element_line(size = 1, colour = "black"),
    ) 

generate_density_plot <- function(data, x_var, output_file) {
  # axis labels
  x_var_label <- ifelse(x_var == "surf_score", "Surface Expression",
                        ifelse(x_var == "ph55_score", "pH 5.5",
                               ifelse(x_var == "ph65_score", "pH 6.5", x_var)))
                                
  # mutation_type mapping
  mutation_type_labels <- c("Syn" = "Synonymous", "Indel" = "Indel", "Mis" = "Missense")

  # order mutation types on plot
  data <- data %>%
    mutate(mutation_type_order = factor(mutation_type, levels = c("Syn", "Indel", "Mis")))

  # create plot
  ggdensity(data, x_var, size = 0.75, color = "mutation_type_order") +
    labs(x = x_var_label, y = "Density") +
    ggtitle(paste(x_var_label, "Density")) +
    scale_y_continuous(breaks = c(0, 0.5, 1), limits=c(0,1)) +  # Set y-axis breaks
    scale_color_manual(values = c("Syn" = "black", "Indel" = "purple", "Mis" = "green"))+
    xlim(-5,5) +
    my_theme + 
    guides(fill = guide_legend(title = ""))
  ggsave(output_file, height = 2, width = 3)
}

# make plots
generate_density_plot(gpr68_scores, "ph55_score", "main/figure4/GPR68_ph55_density.pdf")
generate_density_plot(gpr68_scores, "ph65_score", "main/figure4/GPR68_ph65_density.pdf")
```

surface expression density figure
```{r}
my_theme2 <- theme_minimal() +
    theme(
      legend.position = "none",
      text = element_text(size = 10, color = "black", family = "sans"),
      panel.border = element_rect(color = "black", fill=NA, size=0.5),
      plot.title = element_text(hjust = 0.5),
      panel.grid.major = element_blank(),  
      panel.grid.minor = element_blank(),
      axis.text = element_blank(),
      axis.ticks.x = element_line(size = 0.5, colour = "black"),
      axis.ticks.y = element_line(size = 0.5, colour = "black"),
      axis.ticks.length = unit(0.05, "cm"),
      axis.title = element_blank()
    )

# slightly different function for the surface density plot, largely changing styling
generate_density_plot2 <- function(data, x_var, output_file) {
  # axis labels
  x_var_label <- ifelse(x_var == "surf_score", "Surface Expression",
                        ifelse(x_var == "ph55_score", "pH 5.5",
                               ifelse(x_var == "ph65_score", "pH 6.5", x_var)))
  # mutation_type mapping
  mutation_type_labels <- c("Indel" = "Indel", "Mis" = "Missense", "Syn" = "Synonymous")
    # order mutation types on plot
  data <- data %>%
    mutate(mutation_type_order = factor(mutation_type, levels = c("Syn", "Indel", "Mis")))

  # create plot
  ggdensity(data, x_var, size = 0.3, color = "mutation_type_order") +
    scale_y_continuous(breaks = c(0, 0.5, 1), limits=c(0,1)) +  # Set y-axis breaks
    scale_color_manual(values = c("Syn" = "black", "Indel" = "purple", "Mis" = "green")) +
    my_theme2 +
    guides(fill = guide_legend(title = ""))
  ggsave(output_file, height = 15, width = 30, unit = "mm")
}

# make plots
generate_density_plot2(gpr68_scores, "surf_score", "main/figure4/GPR68_surface_expression_density.pdf")
```
