---
title: "GPR68 Deep Mutational Scan baseline counts"
author: "Matthew Howard, UCSF"
date: "March 11, 2024"
output: html_notebook
---
this notebook contains code used to prepare plots of variant counts in the GPR68 mutational library

some functions for wrangling counts data was adapted or otherwise borrowed from Christian Macdonald's (@UCSF) work as part of https://github.com/odcambc/DIMPLE_manuscript_figures

set working directory, reset to source folder on your system 
```{r}
setwd("/Users/mkh/Library/CloudStorage/Box-Box/WCM Lab/Matt/Lab/Projects/proton_sensors/GPR68_DMS/figures") 
```

source R packages
```{r}
install.packages("tidyverse")
install.packages("ggsci")
install.packages("ggpubr")
install.packages("gridExtra")
install.packages("cowplot")
install.packages("bio3d")
install.packages("tune")
install.packages("MASS")
install.packages("colorspace")
install.packages("readr")
#install.packages("scales")
library(tidyverse)
library(ggsci)
library(ggpubr)
library(gridExtra)
library(cowplot)
library(bio3d)
library(tune)
library(MASS)
library(colorspace)
library(readr)
#library(scales)
```
pull in counts .tsv file
```{r}
counts <- read_tsv("nontreated_R2_counts.tsv")
```

process the hgvs 
```{r}
colnames(counts) <- c("hgvs", "count")
# Parses an HGVS string, outputs a vector with variant info
# Parses an HGVS string, outputs a vector with variant info

parse_hgvs <- function(hgvs_string) {
  variant = ""
  pos = -1
  len = -1
  mutation_type = ""
  
  # WT case, as in enrich2 format
  if (str_detect(hgvs_string, "_wt")) {
    variant = "Z"
    pos = -1
    len = -1
    mutation_type = "X"
  }
  
  # M/S/N
  if (str_detect(hgvs_string, "[A-Z][0-9]+[A-Z]+")) {
    len = 1 
    match = str_match(hgvs_string, "([A-Z])([0-9]+)([A-Z]+)")
    pos = match[3]
    if (match[2] == match[4]) {
      mutation_type = "S"
      variant = match[4]
    } else if (match[4] == 'X') {
      mutation_type = "N"
      variant = match[4]
    } else {
      mutation_type = "M"
      variant = match[4]
    }
  }
  
  # D
  if (str_detect(hgvs_string, ".*del")) {
    mutation_type = "D"
    if (str_detect(hgvs_string, "[A-Z][0-9]+_[A-Z][0-9]+del")) {
      # D_2, D_3
      match = str_match(hgvs_string, "[A-Z]([0-9]+)_[A-Z]([0-9]+)del")
      pos = match[2]
      len = strtoi(match[3]) - strtoi(match[2]) + 1
      variant = paste("D_", as.character(len), sep="")
    } else {
      # D_1
      len = 1
      match = str_match(hgvs_string, "([A-Z])([0-9]+)del")
      pos = match[3]
      variant = "D_1"
    }
  }
  
  # I
  if (str_detect(hgvs_string, ".*ins.*")) {
    mutation_type = "I"
    match = str_match(hgvs_string, "[A-Z]([0-9]+)_[A-Z][0-9]+ins([A-Z]+)")
    len = nchar(match[3])
    pos = match[2]
    variant = paste("I_", as.character(len), sep="")
  }
  return(c(variant, pos, len, mutation_type))
}

## Turn a 3AA HGVS string into a 1AA HGVS string
# Input strings are in form "p.Lys116Lys"
# Output are in form "p.(K116K)"

convert_3AA_hgvs <- function(hgvs_string) {
  if (str_detect(hgvs_string, "_wt")) {
    hgvs_string_1x = "_wt"
  }
  if (str_detect(hgvs_string, "p\\.([A-Za-z]{3})([0-9]+)([A-Za-z]{3})")) {
    match = str_match(hgvs_string, "p\\.([A-Za-z]{3})([0-9]+)([A-Za-z]{3})")
    
    AA1 = toupper(match[2])
    AA2 = toupper(match[4])
    
    AA1 = aa.table[AA1, ]['aa1'][[1]]
    AA2 = aa.table[AA2, ]['aa1'][[1]]
    hgvs_string_1x = paste( 'p.(', AA1, match[3], AA2, ')', sep="")
    
  }
  
  return(hgvs_string_1x)
}

# Parses an HGVS string with 3x AA names, outputs a vector with variant info
# strings are in form "p.Lys116Lys"
# coerce to "p.(K116K)" and call parse_hgvs on it

parse_hgvs_2 <- function(hgvs_string) {
  
  if (str_detect(hgvs_string, "_wt")) {
    hgvs_string_1x = "_wt"
  }
  
  
  #Coerce the HGVS string.
  
  if (str_detect(hgvs_string, "p\\.([A-Za-z]{3})([0-9]+)([A-Za-z]{3})")) {
    match = str_match(hgvs_string, "p\\.([A-Za-z]{3})([0-9]+)([A-Za-z]{3})")
    
    AA1 = toupper(match[2])
    AA2 = toupper(match[4])
    
    AA1 = aa.table[AA1, ]['aa1'][[1]]
    AA2 = aa.table[AA2, ]['aa1'][[1]]
    hgvs_string_1x = paste( 'p.(', AA1, match[3], AA2, ')', sep="")
    
  }
  
  return(parse_hgvs(hgvs_string_1x))
}

# Takes a df with identifiers in HGVS format, outputs
# a df with four additional columns with variant info
process_hgvs_df <- function(df) {
  
  n = nrow(df)
  
  variant_list = setNames(
    as_tibble(
      data.frame(
        rep(NA, n)
      )), 
    c("variants"))
  pos_list = setNames(
    as_tibble(
      data.frame(
        rep(NA, n)
      )), 
    c("pos"))
  len_list = setNames(
    as_tibble(
      data.frame(
        rep(NA, n)
      )), 
    c("len"))
  mutation_type_list = setNames(
    as_tibble(
      data.frame(
        rep(NA, n)
      )), 
    c("mutation_type"))
  
  for(i in 1:n) {
    output = parse_hgvs(df$hgvs[i])
    variant_list[i,] = output[1]
    pos_list[i,] = strtoi(output[2])
    len_list[i,] = output[3]
    mutation_type_list[i,] = output[4]
  }
  
  
  df <- add_column(df, pos_list)
  df <- add_column(df, len_list)
  df <- add_column(df, mutation_type_list)
  df <- add_column(df, variant_list)
  
  return(df)
}

# Takes a df with identifiers in HGVS format with 3-character AA name, outputs
# a df with four additional columns with variant info.
# Note: this is for VatA, which only has substitutions and no indels. Length
# is dropped, and also the positions are already correctly set.

process_hgvs_df_2 <- function(df) {
  
  n = nrow(df)
  
  variant_list = setNames(
    as_tibble(
      data.frame(
        rep(NA, n)
      )), 
    c("variants"))
  pos_list = setNames(
    as_tibble(
      data.frame(
        rep(NA, n)
      )), 
    c("pos"))
  len_list = setNames(
    as_tibble(
      data.frame(
        rep(NA, n)
      )), 
    c("len"))
  mutation_type_list = setNames(
    as_tibble(
      data.frame(
        rep(NA, n)
      )), 
    c("mutation_type"))
  
  for(i in 1:n) {
    output = parse_hgvs_2(df$hgvs[i])
    variant_list[i,] = output[1]
    pos_list[i,] = strtoi(output[2])
    len_list[i,] = output[3]
    mutation_type_list[i,] = output[4]
  }
  df <- add_column(df, mutation_type_list)
  df <- add_column(df, variant_list)
  return(df)
}

counts_processed <- process_hgvs_df(counts)
```

make stacked bar plot 
```{r}
# name a new df of the counts at each pos
counts_by_pos <- counts_processed %>% filter(variants != "Z") %>% group_by(pos) %>% summarise(count = sum(count))

# Create the stacked bar plot
variant_coverage_plot <- ggplot(counts_processed %>% filter(variants != "Z"), aes(x = pos, y = count, fill = variants)) +
  geom_bar(stat = "identity", position = "fill") +
  theme_minimal() +
    labs(title = "Variant coverage per position",
       x = "Position",
       y = "Fraction each variant") +
  theme(legend.position = "right",
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 6),
        axis.text.x = element_text(hjust = 1, size = 6, color = "black"),
        axis.text.y = element_text(size = 6, color = "black"),
        axis.title = element_text(size = 6, color = "black"),
        plot.title = element_text(size = 6, color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.key.height = unit(0.2, "cm"),  
        legend.key.width = unit(0.2, "cm")) +
  scale_x_continuous(limits = c(1,365), expand = c(0.01, 0), breaks = seq(0, max(counts_by_pos$pos), by = 50))
  
coverage_plot <- ggplot(counts_by_pos, aes(x = pos, y = count)) +
  geom_histogram(stat = "identity", bins = 50) +
  labs(title = "Count per position", 
       x = "Position",
       y = "Variant count") +
  scale_y_log10() +
  scale_x_continuous(limits = c(1,365), expand = c(0.01, 0), breaks = seq(0, max(counts_by_pos$pos), by = 50)) +
  theme_minimal() +
  theme(legend.text = element_text(size = 6),
        axis.text.x = element_text(angle = 0, hjust = 0.5, size = 6, color = "black"),
        axis.text.y = element_text(size = 6, color = "black"),
        axis.title = element_text(size = 6, color = "black"),
        plot.title = element_text(size = 6, color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

# use patchwork to arrange
QC_plots <- variant_coverage_plot / (coverage_plot) + plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 6, face = "bold"))

ggsave("supplemental/supp_figure7/QC_plots.png", QC_plots, width = 180, height = 100, units = "mm", dpi = 450)
```